### 2017-03-15 Writing/TrafficAnalysisOn6LoWPAN/conclusion.tex
Moved to this repository.

\section{Conclusion \label{conclusion}}

In this paper, we first reported an implementation issue of 802.15.4 Security in Contiki. Then we presented examples of application of Traffic Analysis over 6LoWPAN network. The first attack shows the content of encrypted ICMP messages could be leak through the packets features such as length and MAC destination. The second attack exploits the time differences in processing a PING Request to distinguish low performance devices used in IoT applications.

%PingProbe Attack
We proposed a new type of side channel attack, PingProbe, in the second half of the paper. This attack exploits the outlier data in PING packets response time to fingerprint different applications running on the target node. We then conducted proof of concept experiments and proved the effectiveness of the attack under our setting.