### 2017-03-20 Writing/TrafficAnalysisOn6LoWPAN/characters.tex
A New Hope

\section{Characters of IoT Traffics}
\begin{itemize}
	\item Less random data (ads, user input, etc).
	\item Different protocol suite (UDP instead TCP, hence no dynamic window.)
	\item Real time applications; hence less noise in timing.
\end{itemize}

Therefore more predictable packet features.

\begin{itemize}
	\item Different approaches: sophisticated techniques may not even required, e.g. machine learning, Mutual Information analysis. Leakage is easily observed.
\end{itemize}

Different scenarios have different leakages, e.g. the channels used in  TSCH could also be a leakage source.

Existing security protocols does not protect these side channel information.