### 2017-03-20 Writing/TrafficAnalysisOn6LoWPAN/countermeasure.tex

\section{Countermeasures}

The constrained environment makes the countermeasures difficult the practice, as they all greatly impacts the performance.

Relatively practical suggestions:
\begin{itemize}
	\item Packet length: Pad to MTU.
	\item Timing: Attempt to response on a fix timeslot.
	\item Protocol: Only enable PING for debugging.
	\item Others: insert dummy packets, etc.
\end{itemize}

However, each application / IoT system needs to tune the countermeasures.

Lower level protections should also be considered. For example physically protect the channel.

More importantly, product designers need to be aware of these potential security and privacy breach.