### 2017-03-20 Writing/TrafficAnalysisOn6LoWPAN/device.tex

\section{Leakage on Device Information}

Different processors result into different response time to the same message; hence the response time can be correlated to the MCU.

\begin{itemize}
	\item PING is a ideal feature in ICMP to launch this attack.
\end{itemize}

The processor can be considered a hint for the adversary to find out the model of target device.