### 2017-03-20 Writing/TrafficAnalysisOn6LoWPAN/plaintext.tex

\section{Leakage on Plaintext}

Examples.

\subsection{ICMP}

Types of ICMP messages can be distinguished by looking at the destination and/or source address together with packet length.

Knowing the ICMP messages may reveal information of the network topology.

\subsection{Sensor Readings with CoAP}

Reading different sensors results into different responding time which can be observed through the packets.