### 2017-04-03 Writing/introduction.tex
Update

\section{Introduction}
{\it
A typical sensor network is built as described in \Cref{Protocols}.

\begin{table}[!h]
	\centering
	\input{tbl/protocols.tex}
	\caption{Protocol stack for sensor networks. (* are optional.)\label{Protocols}}
\end{table}

\begin{itemize}
	\item Security options in 6LoWPAN: 802.15.4 Security and DTLS.
	\item Eavesdropping is easier for 6LoWPAN.
\end{itemize}

We discussed packet features independently in this paper but in practice they can be analysed jointly to reveal more leakage.
}