### 2017-04-05 Writing/appendix.tex
Body draft completed.

\section{Fingerprint Experiment Programs}

\input{tbl/PingProbeExperiments.tex}

\section{Fingerprint Experiment Relative KS-Distances}

\input{tbl/ksdistance.tex}