### 2017-04-06 Writing/iotprotocols.tex
Update

\section{A Typical IoT Protocol Stack}
%A typical sensor network is built as described in \ref{Protocols}.

\begin{table}[!h]
	\centering
	\input{tbl/protocols.tex}
	\caption{Protocol stack for sensor networks. (* are optional.)\label{Protocols}}
\end{table}

%\begin{itemize}
%	\item Security options in 6LoWPAN: 802.15.4 Security and DTLS.
%	\item Eavesdropping is easier for 6LoWPAN.
%\end{itemize}

\subsection{Our experimental network}
%Say something about devices, and choice of OS, etc.