### 2017-04-06 Writing/relatedwork.tex

\subsection{Related Work}
%Write about attacks that exploit packet length and response times in the context of TCP/IP. Reflect on web applications and how the same might happen if more interesting applications might be deployed that involve IoT objects.