### 2017-04-07 Writing/appendix.tex
River publisher ready version 1.0.

\newpage
\subsection*{Fingerprint Experiment Programs}
\begin{table}[!ht]
	\centering
	\input{tbl/PingProbeExperiments.tex}
	\caption{Fingerprint Experiment Programs}
	\label{FingerprintApps}
\end{table}

\subsection*{Fingerprint Experiment Relative KS-Distances}
\begin{sidewaystable}[!ht]
	\footnotesize
	\centering
	\input{tbl/ksdistance.tex}
	\caption{
		Relative KS-Distances of Experimented Fingerprints (Multiplied by 100 for readability. Indexes refer to \Cref{FingerprintApps} 
		Minimum in each row marked as \textbf{bold}.)
	}
	\label{ksdistances}
\end{sidewaystable}